\documentclass[12pt,letterpaper]{report}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[width=20.00in, height=15.00in, left=0.25in, right=0.25in, top=0.25in, bottom=0.25in]{geometry}
\author{Anthony Odenthal, KE7OSN Amateur Extra}
\title{Presentation Notes for the Technician Class Presentation}
\begin{document}
\section*{Preface}
These notes are intended to assist in the giving of the presentation. They should help the presenter cover topics since much of the content of the presentation is designed to touch on topics and have the presenter lecture in more detail. It would also not hurt for the presenter to have read, or have available a copy of the ARRL license manual.

This document is broken into sections that follow the same structure as the presentation.

\section*{Introduction}
Here you should introduce yourself and what amateur radio is. This would be a good time to let your students know how and why you got into amateur radio. Explain how the exam is run, and remind students to not stress.

\section*{FCC rules, descriptions, and definitions}
This covers mostly the administrative issues.

\subsection*{Who's In Charge}
In this part we cover how amateur radio has a good deal of self-regulation and that is done within the framework of the FCC, and that the FCC in turn works within the rules of international treaty, particularly the ITU. Explain what the ITU, FCC, and Frequency Coordinators are. A good example is the 2 meter band, the ARRL band plan carves up the band into small chunks for specific purposes, while the FCC doesn't go into that level of detail.

\subsection*{Part 97}
Next up is more detail on Part 97, here you can explain that for the most part this is where the rules can be found. This presentation emphasizes the noncommercial aspect. Some of the rules are there to protect the amateur way of life, and others are comprises on issues namely stemming from the world wars and the cold war.

\subsection*{Licenses}
License classes, you may go into details about what the novice and advanced were, and how complicated the licensing system was. Also mentioned are types of license, point out in other services the stations and operator licenses are separated. This is also a good time to let advise students that they are not allowed to transmit until they show up in the ULS, but if they upgrade later they can use those upgraded privileges as soon as they pass.

\subsection*{Callsigns}
Go over what makes a valid amateur callsign, a valid us callsign, and a couple other types of calls. Advise that the control operator is the \textit{licensed} operator in charge of the station, it is the control operator's callsign and license that are being used. This presentation very briefly touches on special cases, depending on the students it may or may not be useful to explain these cases in more details.
For an individual working a station whose station license is a lower license than the operator (e.g.\ an extra at a club whose trustee is a general) the ID should be ``station call''/``operator's call''
For a 1x1 the normal callsign should be transmitted at the start or operations (change of frequency), the end of operations (change of frequency), change of normal callsign (change of operators, unless a club call is used), and once every hour. The Id should take the form of ``normal call''/``1x1''
For an operator who has upgraded but the upgrade hasn't shown in the database they should append an ``AG'' for general or ``AE'' for extra to their call, when using the privileges granted to the new class and not the old one. E.g.\ An upgrade to extra from general on 14.175MHz would id as ``callsign''/AE but on 14.300MHz would just ID with their call.

\subsection*{Frequencies}
This is where you should talk about how we share a few of the bands with other users, in mosts cases we are primary, but in a few cases, 30m, 60m,1.25m we have restrictions.

\section*{Station operation; choosing an operating frequency, calling another station, test transmissions, use of minimum power, frequency use, band plans}
This section covers some operation basics, and some things people do with amateur radio. While not in the title this section covers emergency radio.

\subsection*{Emergency Operation}
Remind students that ``normal'' means using a phone, or having another ham use a phone for you. Some places may be procedurally more or less OK with amateurs calling the emergency dispatch directly, nevertheless using spectrum you are not explicitly allowed to when another option exists is a violation of FCC rules, these rules have in fact been enforced. 

For ARES vs RACES the rule difference is that ARES groups can offer help to anyone and there is no national standard for who is in ARES, and how they operate, or who they serve. RACES on the other hand is only active in a declared emergency and only directly serves government agencies. There has been a push to get the two groups cross listed in that hams are ARES until an emergency is declared and then they magically become RACES, this has been adopted in some states.

\subsection*{Nets and messages}
The idea here is to impress upon students that in a case where several amateurs are trying to communicate with each other that without some order. The more people talking the more control is needed to so that people don't talk over each other.

Regardless of the message system, the important thing is that when a message is sent it needs to be sent exactly as given, and that it makes it to its destination exactly as given. The 213 has more widespread usage outside the amateur radio world, NTS is by amateurs for amateurs. 

\subsection*{Radio speak}
Remind students that they don't need to be experts in the jargon, and that they will pick up the important stuff as they go along.

\subsection*{Other practices}
Remind students that they are required to use the minimum power, and that there are power limits, there will be more details at the end.

\section*{Radio wave characteristics, radio and electromagnetic properties, propagation modes}

\subsection*{Electromagnetic waves}
EM waves, they move fast. Voyager is also moving fast so it may be a good idea to look up it's current distance, this can be found on the official NASA Voyager site \url{http://voyager.jpl.nasa.gov/}

\subsection*{Propagation}
Here we talk about how radio waves propagate. To start with try to get students to think of radio wave like light. The key difference is that because of the difference in wavelength the things we think of as causing light to reflect or refract, and what causes radio waves to reflect and refract are different. Different layers of the atmosphere absorb, and refract different radio wavelengths differently. Different parts of the atmosphere react differently to solar energy and will behave differently at night or during a solar storm. 

\subsection*{Propagation Modes}
The different modes that are specifically called out in the test are ``Sporadic E'', ``Tropospheric Ducting'', and ``Knife Edge''. Sporadic E and Tropospheric Ducting work very similar in that signals bounce. Knife Edge causes signals to ``bend'' around a sharp object, like a mountain or ridge.

\subsection*{Polarization}
EM radiation oscillates in straight lines, you can roll the presentation back to the slide with picture of the EM waves. It helps to have an antenna or two handy to help students visualize polarization. 

\section*{Amateur radio practices}
As the preface states the way the information in this section of the question pool is basically a collection of loosely related facts that are nice to know. However there is no cohesive way to present the information. Try to answer student's questions as they come up.

\section*{Electrical principles}
This section is broken into two pieces. The first covers the basics of Ohm's Law, what current is, etc. The second covers the various basic electrical components. Starting here through the rest of the presentation there is a fair amount of technical detail and some students may have issues with 

\section*{How radio works}
Here we start with different modulation modes, CW, AM, FM, and Phase Modulation. Then we show off a few of the basic circuits used to demodulate if you are comfortable with the electronics you can explain things in more detail, but the important point is that SSB and FM are more complex than CW and AM. Later we look at some higher level schemes built on top of those modulation methods. If you are familiar with the 7-layer OSI model you can discus that here.

\section*{Antennas and feedline}
No matter how much power one runs it is important to have a good antenna. A poor antenna system can waste all of the power as heat, at low power almost now energy goes out radio waves. At high power levels arcing, or even melting can happen. A simple quarter wave vertical demonstrates the very basics. The progression through dipole, to the Yagi shows how many of the complex antennas are built from simple building blocks. 

For feedline the two differences to look at are how signals move though it and how a signal is protected from interference. Waveguide conducts radio waves through them, while cables move electric current that must transition between radio wave and current at an antenna. Balanced intends to use the fact that all wires receive the same interference, but that they don't carry the same signal. Coax on the other hand relies on a grounded shield to protect the center conductor from noise. The larger the ``wire'' the less resistance, meaning lower loss and the ability to safely carry more power.

The impedance section is a lot of text and it may help to simplify things down to match impedance, reduce the SWR and things work better. There are many connector options and it is important to know that there is no one best fit for all cases, all have their uses. 

\section*{Power and Safety}
In addition to the examples in the slides that the voltage at the end(s) of an antenna can get very high. It is highly recommended that people go out of their way to prevent shocks, falls or other threats to life. Grounding of electrical systems including radios is not only important for safety, but properly grounded systems can greatly improve operating conditions.

Most of the rest of the safety items should be common sense, and the power limits are a matter of just look it up.

\section*{Best practices}
As the slide indicates these tidbits are here purely as a PSA. This is a good place to simulate a QSO or two to illustrate how one actually talks to another ham.

\end{document}
